---
title: "SEC_BIAS_HEISMAN"
author: "Nathaniel Ginck"
date: "10/24/2023"
output: html_document
---
Install packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
installed.packages("MASS")
library(MASS)
install.packages("regclass")
library(regclass)
install.packages("corrplot")
library(corrplot)
install.packages("ggplot2")
library(ggplot2)
install.packages("readr")
library(readr)
```

Read in data
```{r}
QB_Data <- read.csv("Player Data 2000 to 2022.csv")
```

Tidy data
```{r}
# first remove 2000,2001, 2002 as there is no available strength of conference available
QB_Data_Tidy <- QB_Data[!(QB_Data$Year %in% c(2000,2001,2002)), ]

# next remove position, player name, school, and school adjusted
QB_Data_Tidy <- QB_Data_Tidy[, -c(2,3,4,5)]

# remove players that received 0 points
QB_Data_Tidy <- QB_Data_Tidy[QB_Data_Tidy$Heisman.Points != 0, ]

# Map Conference to binary (SEC = 1, other = 0)
QB_Data_Tidy$Conf <- ifelse(grepl("SEC", QB_Data_Tidy$Conf), 1, 0)

# remove extraneous columns in Data set
QB_Data_Tidy <- QB_Data_Tidy[, -c(20,21,22,23,24,25)]

#remove unused statistics from data set
QB_Data_Tidy <- QB_Data_Tidy[, -c(1,3,5,6,7,8,9,10,13,14,15,16,18)]

# Add a column: natural log of Heisman Points
QB_Data_Tidy$LnHeismanPoints <- log(QB_Data_Tidy$Heisman.Points)

write_csv(QB_Data_Tidy, "Tidied_QB_Data.csv")


```

Descriptive Data:
Histogram of data
```{r}
png("HistogramHeismanPts.png")
# histogram of points
Histogram = hist(QB_Data_Tidy$Heisman.Points, main = "Histogram of Quarterback Heisman Points", xlab = "Quarterback Heisman Points")
```
Create Boxplot
```{r}
png("BoxplotHeismanPts.png")
# boxplot of points
boxplot(QB_Data_Tidy$Heisman.Points, data = QB_Data_Tidy, main = "Boxplot of Quarterback Heisman Points", horizontal = TRUE)
```
Check homoscedasticity:

Create models
```{r}
ModelPassTD = lm(QB_Data_Tidy$Heisman.Points ~ QB_Data_Tidy$Pass.TD)
ModelRushTD = lm(QB_Data_Tidy$Heisman.Points ~ QB_Data_Tidy$Rush.TD)
ModelInt = lm(QB_Data_Tidy$Heisman.Points ~ QB_Data_Tidy$Pass.Int)
```

Create Scatterplots
```{r}
png("ScatterPassTD.png")
plot(QB_Data_Tidy$Pass.TD, QB_Data_Tidy$Heisman.Points, pch = 19, xlab = "Passing Touchdowns", ylab = "Heisman Points")
abline(ModelPassTD)
```
Create Scatterplot
```{r}
png("ScatterRushRD.png")
plot(QB_Data_Tidy$Rush.TD, QB_Data_Tidy$Heisman.Points, pch = 19, xlab = "Rushing Touchdowns", ylab = "Heisman Points")
abline(ModelRushTD)
```
Create Scatterplot
```{r}
png("ScatterInt.png")
plot(QB_Data_Tidy$Pass.Int, QB_Data_Tidy$Heisman.Points, pch = 19, xlab = "Interceptions", ylab = "Heisman Points")
abline(ModelInt)
```


Residual plots:
```{r}
png("RushResid.png")
plot(predict(ModelRushTD), residuals(ModelRushTD), pch = 19, ylab = "Residuals", xlab = "Rushing Touchdowns")
abline(h = 0)
```

Residual plot
```{r}
png("PassResid.png")
plot(predict(ModelPassTD), residuals(ModelPassTD), pch = 19, ylab = "Residuals", xlab = "Passing Touchdowns")
abline(h = 0)

```

Residual plot
```{r}
png("IntResid.png")
plot(predict(ModelInt), residuals(ModelInt), pch = 19, ylab = "Residuals", xlab = "Interceptions")
abline(h = 0)
```


Adjusted histogram with transformation of response variable
```{r}
# Obtain new histogram
png("LnPointsHist.png")
Histogram = hist(QB_Data_Tidy$LnHeismanPoints, main = "Histogram of Natural Log of Quarterback Heisman Points", xlab = "Natural Log of Quarterback Heisman Points")

```

New boxplot
```{r}
# Obtain new boxplot using LN
png("LnBoxPlot.png")
boxplot(QB_Data_Tidy$LnHeismanPoints, data = QB_Data_Tidy, main = "Boxplot of Natural Log of Quarterback Heisman Points", horizontal = TRUE)

```

Summary
```{r}
# Obtain summary
summary(QB_Data_Tidy$LnHeismanPoints)
```

Look at fit of predictors in model
```{r}
# Create simple single variable models for three predictors
LnModelPassTD = lm(QB_Data_Tidy$LnHeismanPoints ~ QB_Data_Tidy$Pass.TD)
LnModelRushTD = lm(QB_Data_Tidy$LnHeismanPoints ~ QB_Data_Tidy$Rush.TD)
LnModelInt = lm(QB_Data_Tidy$LnHeismanPoints ~ QB_Data_Tidy$Pass.Int)
```

Plot Scatterplots with line of best fit
```{r}
png("LnPassTD.png")
plot(QB_Data_Tidy$Pass.TD, QB_Data_Tidy$LnHeismanPoints, pch =19, ylab = "Natural Log of Heisman Points", xlab = "Passing Touchdowns")
abline(LnModelPassTD)
```
```{r}
png("LnRushTD.png")
plot(QB_Data_Tidy$Rush.TD, QB_Data_Tidy$LnHeismanPoints, pch =19, ylab = "Natural Log of Heisman Points", xlab = "Rushing Touchdowns")
abline(LnModelRushTD)
```
```{r}
png("LnInt.png")
plot(QB_Data_Tidy$Pass.Int, QB_Data_Tidy$LnHeismanPoints, pch =19, ylab = "Natural Log of Heisman Points", xlab = "Interceptions")
abline(LnModelInt)
```

Plot Residual plots
```{r}
png("LnRushResid.png")
plot(predict(LnModelRushTD), residuals(LnModelRushTD), pch =19, ylab = "Residuals", xlab = "Rushing Touchdowns")
abline(h = 0)
```
```{r}
png("LnPassResid.png")
plot(predict(LnModelPassTD), residuals(LnModelPassTD), pch =19, ylab = "Residuals", xlab = "Passing Touchdowns")
abline(h = 0)
```
```{r}
png("LnIntResid.png")
plot(predict(LnModelInt), residuals(LnModelInt), pch =19, ylab = "Residuals", xlab = "Interceptions")
abline(h = 0)
```

Schedule Strength:
Create model and scatterplot
```{r}
# Confirming that linear model is appropriate
ScheduleStrengthModel = lm(LnHeismanPoints ~ (Schedule.Strength), data = QB_Data_Tidy)
plot(QB_Data_Tidy$Schedule.Strength, QB_Data_Tidy$LnHeismanPoints, pch = 19, xlab = "Schedule Strength", ylab = "Natural Log of Heisman Points")
abline(ScheduleStrengthModel)
summary(ScheduleStrengthModel)

```
Residual plot for SOS:
Confirming homoscedasticity is not violated
```{r}
png("LnSOSResid.png")
plot(predict(ScheduleStrengthModel), residuals(ScheduleStrengthModel), pch =19, ylab = "Residuals", xlab = "Schedule Strength")
abline(h = 0)

```


Inferential Data

Models
```{r}
# Base Model
Base_Model <- lm(QB_Data_Tidy$LnHeismanPoints ~ QB_Data_Tidy$Rush.TD + QB_Data_Tidy$Pass.TD + QB_Data_Tidy$Pass.Int)

summary(Base_Model)

Model_Conf <- lm(LnHeismanPoints ~ Rush.TD + Pass.TD + Pass.Int + Conf, data = QB_Data_Tidy)

summary(Model_Conf)
```
Testing if collinearity is a problem
```{r}
# calculate VIF
VIF(Model_Conf)

# calc and display correlations
sprintf("The correlation between SEC and Rushing Touchdowns is: %f", cor(QB_Data_Tidy$Conf, QB_Data_Tidy$Rush.TD))
sprintf("The correlation between SEC and Passing Touchdowns is: %f", cor(QB_Data_Tidy$Conf, QB_Data_Tidy$Pass.TD))
sprintf("The correlation between SEC and Interceptions is: %f", cor(QB_Data_Tidy$Conf, QB_Data_Tidy$Pass.Int))
```

Test interaction
```{r}
# create interaction model
InteractionModel = lm(LnHeismanPoints ~ Conf*Schedule.Strength, data = QB_Data_Tidy)
summary(InteractionModel)

# show anova summary
anova(InteractionModel)

```
